---
title: Ubuntu Install Notes / Linux Documentation
author: Ryan Nash
output: pdf_document
header-includes:
  - \hypersetup{colorlinks=true, linkcolor=black}
---

\tableofcontents

#USB Install
* Boot into Xubuntu live usb
* Format hdd using `parted` or `gparted`
* Run through Xubuntu install process
* Reboot

#Newly Installed System
* Update packages
    + `sudo apt update`, `sudo apt upgrade`, `sudo apt dist-upgrade`
* Reboot
    
#Install Drivers

##GPU
* Install Nvidia proprietary driver from *Software & Updates* in **Settings**
* Disable Nvidia GPU in `nvidia-settings` for better battery life

##X11
* Copy `xorg.conf.d` folder from dotfiles to `/etc/X11/`

##Battery Control
* Install `tlp` for better battery management
    + `sudo apt install tlp tlp-rdw acpi-call-dkms`
    + `sudo tlp start`

##Fan Control 
* Install `thinkfan` for better fan control
    + sudo bash
    + apt install thinkfan
    + echo "options thinkpad_acpi fan_control=1" > /etc/modprobe.d/thinkfan.conf
    + sh -c 'echo coretemp >> etc/modules'
    + modprobe thinkpad_acpi && modprobe coretemp
    + nano -w /etc/default/thinkfan
        + add `START=yes` to second line (below comment)
    + copy `thinkfan.conf` from dotfiles to `/etc/`
    + systemctl enable thinkfan.service
    + reboot

## Network
* Sign-in to *Firefox Sync*
* Disable WebRTC
    + Type `about:config` in *URL bar*
    + Search for `media.peerconnection.enabled`
    + Toggle value to `false`
        + To reverse it, simply toggle back 
* Fix right-click context menu when using i3
    + Go to `about:config` again
    + Change `ui.context_menus.after_mouseup` from **false** to **true**

### VPN
* Install **Mullvad** configuration file for **OpenVPN**
    + `sudo apt-get install openvpn network-manager-openvpn network-manager-openvpn-gnome`
    + Download config file from `https://mullvad.net/en/download/config/`
    + Open the downloaded file and remove everything from `<crl-verify>` to `</crl-verify>` (including them) 
    + Import a saved VPN configuration using **Network Manager** 
    + Edit the connection
        + Enter **Mullvad** account number in the `username` field
        + Enter `m` in the `password` field 
    + `sudo service network-manager restart`
    + Click on *Network Icon* and select created VPN
    + *Mullvad with OpenVPN currently has DNS leaks after doing the above steps. Figure out a fix!*

# Ricing

## Dotfiles
* Copy `.Xresources` to *user home folder*
    + `xrdb .Xresources`
* Copy `compton.conf` to `~/.config/`

## Fonts
* Create `~/.fonts directory`
* Copy `.ttf` files to separate folders within `.fonts`
    + **System San Francisco**
    + **Adobe Source Code Pro**
    + **FontAwesome**
* Set **Source Code Pro** as the font in:
    + `xfce4-terminal`
    + **XFCE Settings > Appearance > Fonts > Default monospace font**
* Set **SFNS Display**/San Francisco Display as the font in:
    + **XFCE Settings > Appearance > Fonts**
    + **XFCE Settings > Window Manager**
* Customise `xfce4-terminal` using its *preference* GUI    

## GTK Themes & Icons
* Install `arc-theme`
* Install `numix-solarized` theme
* Install `macos-sierra` theme
    + and dependencies 
* Install `papirus-icon-theme`
* Install MacOS icons
* For **XFCE** set theme/icons in:
    + **XFCE Settings**
    + **Window Manager**
* For **i3** set theme/icons in:
    + lxappearance (`sudo apt install lxappearance`)
* Setting wallpaper in XFCE will also set it for i3

## i3wm
* Install the **i3 window manager**
    + `sudo apt install i3-suckless-tools i3blocks i3lock`
* Copy config file from dotfiles to `~/.config/i3/`

## i3-gaps
* Install **i3-gaps** dependencies

```{r, eval = FALSE}
$ sudo apt install \
libxcb1-dev libxcb-keysyms1-dev libpango1.0-dev \
libxcb-util0-dev libxcb-icccm4-dev libyajl-dev \
libstartup-notification0-dev libxcb-randr0-dev \
libev-dev libxcb-cursor-dev libxcb-xinerama0-dev \
libxcb-xkb-dev libxkbcommon-dev libxkbcommon-x11-dev \
autoconf libxcb-xrm0 libxcb-xrm-dev automake
```
* Clone the repository
    + `git clone https://www.github.com/Airblader/i3 /tmp/i3-gaps`
    + `cd /tmp/i3-gapsr
* Compile & install
```{r, eval = FALSE}
$ autoreconf --force --install
$ rm -rf build/
$ mkdir -p build && cd build/
```
* Disabling sanitizers is important for release versions!
* The prefix and sysconfdir are, obviously, dependent on the distribution.

```{r, eval = FALSE}
$ ../configure --prefix=/usr --sysconfdir=/etc --disable-sanitizers
$ make
$ sudo make install
```
* After installing i3 and i3-gaps, `dunst` took precedence over the nicer looking `xfce4-notify-d`, so I remove it
    + `sudo apt remove dunst`
    + `sudo apt autoremove`
* Install **git**, `sudo apt install git`

## i3blocks
* Custom scripts: (currently located in: `~/.config/i3/i3blocks/`)
    + IntBat
    + ExtBat
* These scripts need to be made executable in order for them to appear in i3blocks bar
    + `sudo chmod +x ScriptName`

## Powerline
* `sudo pip install powerline-status`  
* `sudo apt install powerline`
* `sudo pip install powerline-gitstatus`

* Add to `.bashrc` and `.vimrc`
* Configure colors using `config.json` 

### Powerline-gitstatus
*

## Vim
* Install `vim-gtk`
    + `sudo apt install vim-gtk`
* Copy `.vimrc` config file and `.vim` folder from dotfiles to `~/`
* Set vim as default editor
    + `sudo update-alternatives –config editor`
* To use my .vimrc while editing root owned files `sudo -E vi ~/.vimrc`
Edit vim solarized colorscheme

### Vim Plugins
* Create ~/.vim/pack/plugins/start/
    + This is the folder to install plugins in
* Example:
    + `cd ~/.vim/pack/plugins/start`
    + `sudo git clone https://github.com/tpope/vim-sensible.git`


## LaTeX
* Install Ubuntu LaTeX package
    + `sudo apt install texlive-full`
* After installing LaTeX, FontAwesome icons stop rendering in i3
    + To fix this, create the directory `~/.config/fontconfig/conf.d/`
    + Copy `/etc/fonts/conf.d/60-latin.conf` to the above folder
    + Edit all `<family>xxxxxx</family>` sections to `<family>FontAwesome</family>` 
    + Except **heading lines** (e.g. serif)
* Install `mupdf` and `xdotool` packages for pdf viewing
 
## R Markdown
* Install R mardown and required packages
    + `sudo apt install r-base r-base-dev pandoc pandoc-citeproc`
* Run R session
    + `sudo R`
    + Install `rmarkdown`, `install.packages("rmarkdown")`
    + Exit R

## SC-IM (CLI-based spreadsheet editor)
* Install required dependencies and `libxlsxwriter`
```{r, eval = FALSE}
$ sudo apt-get install bison libncurses5-dev libncursesw5-dev libxml2-dev libzip-dev
$ git clone https://github.com/jmcnamara/libxlsxwriter.git
$ cd libxlsxwriter/
$ make
$ sudo make install
```
* Refresh dynamic link cache
    + `sudo ldconfig`
* Download (git clone) and compile SCI-IM
```{r, eval = FALSE}
$ cd ..
$ git clone https://github.com/andmarti1424/sc-im.git
$ cd sc-im/src
$ make
$ sudo make install
```
* Run with `sc-im` command, exit same way as **Vim**

# Music Player

## Mopidy with Spotify using ncmpcpp
* Add mopidy archive gpg key 
    + `wget -q -O - https://apt.mopidy.com/mopidy.gpg | sudo apt-key add -`
* Add mopidy to sources 
    + `sudo wget -q -O /etc/apt/sources.list.d/mopidy.list https://apt.mopidy.com/stretch.list`
* `sudo apt update`
* Install `mopidy` and `mopidy-spotify` packages
* Install `ncmpcpp` package
* Install `mpc` package (used for outputting mopidy now playing to bars/i3blocks)
* Copy mopidy folder from dotfiles to `~/.config/`
* Copy .ncmpcpp folder from dotfiles to `~/.`
* Run mopidy in terminal foreground for first run `mopidy`
* Run `ncmpcpp` and test out Spotify
* Install `socat` package to watch UDP port for visualiser
* End terminal process, run my **Visualiser batch script**
* *Note that my Visualiser script must be run on each boot*

### Getting Spotify playlists to work
* Kill all instances of mopidy, then remove installed `mopidy-spotify` package
* `sudo apt-get remove mopidy-spotify`
    + `sudo apt autoremove`
* Install mopidy-spotify dependencies
    + `sudo apt install libspotify12 python-cffi python-ply python-pycparser python-spotify`
* Clone mopidy-spotify from GitHub repo
    + `git clone https://github.com/BlackLight/mopidy-spotify.git /tmp/mopidy-spotify`
*Go to cloned directory
    + `cd mopidy-spotify`
* Switch to `fix/incompatible_playlists` branch
    + `sudo git checkout fix/incompatible_playlists`
* Install cloned-mopidy spotify
    + `sudo python2 setup.py build install`

### Visualiser Script Breakdown    
* create fifo file for visualiser `mkfifo /tmp/mpd.fifo`
* run `while :; do socat -d -d -T 1 -u UDP4-LISTEN:5555 OPEN:/tmp/mpd.fifo; done &`
* notice the '&' at the end, this lets the command run in the background, freeing up terminal space

## Misc Required Programs
* `xbacklight` (controls brightness in i3)
* `redshift` (warmer colors at night)
* `rofi` (used by i3 to launch programs instead of `dmenu`)
    + customised using `.Xresources` file
* `compton` (window compositor I use with i3)
    + Customised with `~/.config/compton.conf` 
    + Currently I only have an opacity rule
* Timeshift
    + `sudo add-apt-repository -y ppa:teejee2008/ppa`
    + `sudo apt update`
    + `sudo apt install timeshift`
* **VSCode** (GUI text editor for use with XFCE)
* Transmission
    + In *preferences* set to `Stop seeding at ratio:` to 0
* Pip
    + `sudo apt install python-pip
* S-Tui, install using Pip
    + `sudo pip install s-tui`
* CalCurse (terminal calender/to list, *install package* `calcurse`)
* feh (background setter and image viewer for i3)

# Aliases
* Located in `~/.bashrc`
    + Syntax is `alias **name**='path to script/command name'`

# Virtualisation
* Install `libvirt-bin` and `virt-manager`
* Add `intel-iommu=on` to `GRUB_CMDLINE_LINUX_DEFAULT` between quotes
* Run `sudo update-grub` then reboot

# Bumblebee
* Install `sudo apt-get install bumblebee bumblebee-nvidia primus linux-headers-generic`
